---
title: "11_import-clean"
output: html_notebook
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Motivating Question

There is a lot of data regarding how people get injured. I'm really curious if the hospital emergency department could be more efficient if it could predict the types of injuries and individuals arriving on any given day. I think there is value in both staffing, supplies, etc. So here we will look at some data about hospital emergency room events and try to use it to generate a model that can be used to predict incoming patients.

# Libraries
```{r Libraries}
library(tidyverse)
library(janitor)
library(fs)
library(lubridate)
library(data.table)
library(assertr)
#install.packages("naniar")
library(naniar)
```

# NEISS Injury Data

# Imports Cleaning and Organizing

First we are going to want to pull in the data, then we will clean it up and get it in a state where we can start applying features.

## Importing

We need to import the raw NEISS data, as well as a dataframe with how the data is encoded.

### Code Indecises

Much of the data is encoded, in order to make it understandable to a human we are going to use the following data to join readable names to the data set. The data is organized below.

```{r code references import}
coderef <- read_delim("data/code-ref/neiss-coderef.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE) %>% clean_names()
```
## Organizing

Unfortunately the coding is a single table. I needed to break it up into decoders for each column of data in the dataset.
This is done here:

```{r Code References Splitting}
coderefsplit <- coderef %>%
  group_by(format_name) %>%
  group_split() %>%
  set_names(unique(coderef$format_name))
  
age <- coderefsplit$AGELTTWO %>% 
  select(starting_value = starting_value_for_format, ending_value = ending_value_for_format, label = format_value_label)

bodypart <- coderefsplit$BDYPT %>% 
  select(code = starting_value_for_format, desc = format_value_label)

diagnosis <- coderefsplit$DIAG %>% 
  select(code = starting_value_for_format, desc = format_value_label)

disposition <- coderefsplit$DISP %>% 
  select(code = starting_value_for_format, desc = format_value_label)

fire_involve <- coderefsplit$FIRE %>% 
  select(code = starting_value_for_format, fire_involve = format_value_label)

gender <- coderefsplit$GENDER %>% 
  select(code = starting_value_for_format, gender = format_value_label)

location <- coderefsplit$LOC %>% 
  select(code = starting_value_for_format, location = format_value_label)

race <- coderefsplit$RACE %>% 
  select(code = starting_value_for_format, race = format_value_label)

product_ref <- coderefsplit$PROD %>% 
  select(code = starting_value_for_format, product = format_value_label)
```

I identified an issue with the data. The following is a function that fixes the issue, it is called during the import to correct for some data that gets split between to rows.


```{r function to fix broken lines}
clean_neiss <- function(df){
  for(i in seq_along(nrow(df))) { 
  if(is.na(as.numeric(df[[i,1]]))){
    df$narrative_1[i-1] <- paste0(df$narrative_1[[i-1]]," ", df$cpsc_case_number[[i]])
    df$narrative_2[i - 1] <- df$treatment_date[i]
    df$psu[i - 1] <- df$sex[i]
    df$weight[i - 1] <- df$race[i]
    #neiss <- neiss[-c(i),] #Why did this not work?
    df <- subset(df, !df[i,])
  }
}

df <- subset(df, !is.na(as.numeric(df$cpsc_case_number)))

}
```
Solving it this way lost us a bit of Narrative_1 and the Stratum value, though I am not sure what Stratum even is. We will drop this column after import.



## Importing the NEISS data. This is the raw data from NEISS
```{r Raw Data Import}

neiss1999 <- read_delim("data/neiss1999.tsv", 
    "\t", escape_double = FALSE, col_types = cols(Age = col_character(), 
        CPSC_Case_Number = col_character(), 
        Other_Diagnosis = col_double()), 
    na = "NA", trim_ws = TRUE) %>% 
    clean_names() %>% #clean the names of the data
    clean_neiss() #This function is below 
  
```
Ignore this, error, it is working as intended, this is an artifact of the cleaning function.

Dropping the columns we will not need.
```{r drop unecessary columns}
neiss1999 <- neiss1999 %>% 
  select(-c(narrative_1, narrative_2, stratum, psu, other_race))
```


There were a couple of rogue data points in the product_2 column these are delt with here.
Additionally we convert the treatment_date column into an actual date time column. We wait to do that here as the problem described above (in the clean_neiss function) required importing the column as a character type.
```{r Cleaning Data}
#Solves issue of the 5 0's mentioned below
neiss1999 <- neiss1999 %>% replace_with_na(replace = list(product_2 = 0))

#Converting Date to a date time, waited until here because there was bad data in this field up until this point
neiss1999 <- neiss1999 %>% 
  mutate(treatment_date = mdy(treatment_date))
```

# Import Asserts - This chunk runs import asserts on the data. It uses the encoding data from the first document imported and confirms that all of the data in the data set meets these required values.
```{r Assertions, error=TRUE}

neiss1999 %>%
  assert(is_uniq(neiss1999$cpsc_case_number))
neiss1999 %>%
  assert(in_set(age$starting_value, 2:120), age)
neiss1999 %>%
    assert(in_set(c(0:3)), sex)
neiss1999 %>%
  assert(in_set(race$code), race)
neiss1999 %>%
  assert(in_set(bodypart$code), body_part)
neiss1999 %>%
  assert(in_set(diagnosis$code), diagnosis)
neiss1999 %>%
  assert(in_set(disposition$code), disposition)
neiss1999 %>%
  assert(in_set(location$code), location) 
neiss1999 %>%
  assert(in_set(fire_involve$code), fire_involvement)
neiss1999 %>%
  assert(in_set(product_ref$code), product_1)
neiss1999 %>%
  assert(in_set(product_ref$code), product_2)
neiss1999 %>%
  assert(within_bounds(0,200), weight)

```