---
title: "30_feature_engineering"
output: html_notebook
---

# Feature 1 - Grouping patients my standard medical age groups

So I did some research on standard medical age groups...There is a lot of information out there. I'm using here what the following paper suggested was the past standard for medical age groups. These can be updated as desired, these are used here as a baseline.

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3825015/

This feature should help in making sense of how the data changes by age. If we know teenagers often do something it will be easier to see than to have to model on a selection of years. It will now be built in.

```{r Bin Patients Ages}
agebinner <- function(df, var) {
  df %>%
    mutate(age_group = findInterval({{var}}, c(2, 6, 14, 19, 34, 49, 65, 79, 99, 120, 201, 223), rightmost.closed = TRUE))
}
```

## Testing Age Binning Function
```{r Age Bin Test}
neiss1999 %>% 
  agebinner(age) %>% 
  select(age_group, everything()) %>% 
  arrange(-age_group) %>% 
  assert(in_set(1:11)) #the function puts ages in to 11 bins, this confirms that it is one of those 11.
```
The Age binning is working as intended.
NOTE: Bins 10 and 11 correspond to infants between 0-12 and 12-24 months. 


# Feature 2 - Adding Seasonal Information

There were a few options here as far as adding a season column:
https://en.wikipedia.org/wiki/Season

My initial thought was to use the solstice/equinoxes. This is a little challanging as the dates change every year. But, as the dataset includes information from all over the US being incredibly technical about season dates seems unnecessary and cumbersome, so we will use the "Meteorological" definition of seasons as found in the above wikipedia page. These are easier to work with as they are standard year to year.


Seasons affect whether people are outdoors or indoors, or maybe it is too cold to drive to the hospital, I believe this will play in to trends in hospital visits.
```{r Feature 4 Assigning Seasons}
seasons <- function(x){
  if(x %in% 2:4) return("Spring")
  if(x %in% 5:7) return("Summer")
  if(x %in% 8:10) return("Fall")
  if(x %in% c(11,12,1)) return("Winter")
}

assign_season <- function(df, var){
  df %>% 
  mutate(season = sapply(month({{var}}), seasons))
}
```

## Testing Feature 4, seasonal identifier
```{r Testing Assigned Season Function}
neiss1999 %>% 
  assign_season(treatment_date) %>% 
  select(season, everything()) %>% 
  assert(in_set(c("Spring", "Summer", "Fall", "Winter")))
```

===================================================
Add new feature here: Feature 3 or Feature 4
===================================================
# *Feature 3* - Appending a Daily count to each observation
Very important to what I'm trying to explore is the amount of visits per day, this feature is going to add a column which attaches the totals by day to each date.

```{r}
daily_count <- function(df, var) {
  df %>%
    group_by({{var}}) %>%
    mutate(daily_visits = n())
}
```


# *Feature 4* - Add the weekday to the data
Weekday information is always useful when trying to figure out when things happen. People behave differently depending on where they are in the week. Preliminary exploration of the data found a significant drop in visits on Fridays, and significant increases on Saturday and Sunday.


